Example Manuscript Template for a Data Analysis Project

### 2024-03-07 R/eda-code/eda_superstore.qmd
Project Part 2 - Rev

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

# Setup
Required packages are loaded.
```{r}
#load needed packages. 
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Data is loaded.

@@

# Data exploration through figures

Histograms were plotted for the continuous outcomes. The data contains individual observations for sales of different stocks in a single invoice. As this project intends to study customer spending, Invoice totals are calculated for purchases by each customer at a time to visualize distribution of customer spending each time.The histogram is heavily right-skewed with higher frequencies at less than 500.

```{r}
# Histogram of Invoice values
## First calculate total invoice value per invoice
invoice_values <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalInvoiceValue = sum(Sales, na.rm = TRUE))

##Create histogram
p1 <- ggplot(invoice_values, aes(x = TotalInvoiceValue)) +
  geom_histogram(binwidth = 50, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Total Invoice Values",
       x = "Total Invoice Value", 
       y = "Frequency") +
  theme_minimal()
plot(p1)

figure_file = here("results","figures","Sales_distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Similarly, histogram plot was created for total Quantity of purchase per invoice which is also heavily right-skewed.

```{r}
## First calculate total invoice quantity per invoice
invoice_quantity <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalInvoiceQuantity = sum(Quantity, na.rm = TRUE))

p2 <- ggplot(invoice_quantity, aes(x = TotalInvoiceQuantity)) +
  geom_histogram(binwidth = 50, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Total Invoice Quantity",
       x = "Total Invoice Quantity", 
       y = "Frequency") +
  theme_minimal()
plot(p2)

figure_file = here("results","figures","Quantity_distribution.png")
ggsave(filename = figure_file, plot=p2)
```

Histogram is also plotted for price range of the products.Most of the products fall below the price of 5 sterling pounds.

@@

p3 <- ggplot(mydata, aes(x = UnitPrice)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Unit Price",
       x = "Unit Price", 
       y = "Frequency") +
  theme_minimal()
plot(p3)

figure_file = here("results","figures","UnitPrice_distribution.png")
ggsave(filename = figure_file, plot=p3)
```

Density plot is created for total purchase by each Customer during the period of the data. The average total purchase amount appears to be less than 5000 sterling pounds during that period.

```{r}
# Aggregate sales by CustomerID
sales_per_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  arrange(desc(TotalSales))

#Density Plot of Sales per Customer
# Assuming sales_per_customer is already calculated
p4 <- ggplot(sales_per_customer, aes(x = TotalSales)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density of Sales Across Customers", x = "Total Sales", y = "Density")

plot(p4)

figure_file = here("results","figures","Customer_AggregateSales.png")
ggsave(filename = figure_file, plot=p4)

@@

Next, histogram of purchase frequency of each customer is created by grouping each customer by invoice no.The frequency appears to be less than 5 times for most of the customers.

```{r}
# Calculate purchase frequency per customer
purchase_frequency <- mydata %>%
  distinct(CustomerID, InvoiceNo) %>% # Ensure unique InvoiceNo per CustomerID
  count(CustomerID, name = "PurchaseFrequency") # Count the number of invoices per customer
library(ggplot2)

# Histogram of purchase frequencies
p5 <- ggplot(purchase_frequency, aes(x = PurchaseFrequency)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Customer Purchase Frequencies",
       x = "Purchase Frequency", 
       y = "Count of Customers") +
  theme_minimal()
plot(p5)
figure_file = here("results","figures","Customer_purchasefreq.png")
ggsave(filename = figure_file, plot=p5)

Next, total purchase by a customer was plotted against purchase frequency.A few outliers were noted which needs to be addressed while model building.

```{r}
# Calculate purchase frequency and total invoice value per customer
customer_stats <- mydata %>%
  group_by(CustomerID) %>%
  summarise(PurchaseFrequency = n_distinct(InvoiceNo),
            TotalInvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p6 <- ggplot(customer_stats, aes(x = PurchaseFrequency, y = TotalInvoiceValue)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Purchase Frequency",
       y = "Total Invoice Value") +
  theme_minimal()
plot(p6)

figure_file = here("results","figures","Sales_trend.png")
ggsave(filename = figure_file, plot=p6)

Bar chart of Sales Distribution by Country

>>>>

---
title: "Exploratory Analysis of Superstore"
author: "Malika Dhakhwa"
date: "2024-02-23"
output: html_document
---

# Setup
Required packages are loaded.
```{r}

#| message: false
#| warning: false
#load needed packages. 
library(here) #for data loading/saving
library(dplyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(ggplot2) # for visualization
library(tibble) # For rownames_to_column
library(gt) # for table
library(gridExtra)
library(webshot2) #Capture as an Image the gttable before converting it to pdf
```

Data is loaded using here function.

@@

# Descriptive Statistics tables

It is important to note that this data accounts sales of a particular item at a time as an observation leading different items sold in the same Invoice to appear as separate observations. As such most of the Invoice numbers appear several times in different observations of the data. A meaningful analysis of customer spending pattern require the Invoice Sales Amount, total purchase by a Customer during a period, frequency of purchase by a customer during a period or a combination of those. A table is created which provides the descriptive statistics of sales by Invoice value, Customer purchase and frequency of customer purchase. This table reveals that there were 1479 transactions to 414 different customers during 2010-12-01 and	2011-12-09.

```{r}

# Aggregate total sales by InvoiceNo for calculating mean, standard deviation, minimum and maximum
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>% # Ensure we ungroup before further summarisation
  summarise(
    Count = n(),
    Particulars = "Sales by Invoice",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

# Aggregate total sales by CustomerID to calculate mean, standard deviation, minimum and maximum
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Particulars = "Cumulative Sales to Individual Customers",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

# Frequency of sales per CustomerID (i.e., the count of unique InvoiceNo per CustomerID)for calculating mean, standard deviation, minimum and maximum

sales_frequency_per_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(FrequencyOfSales = n_distinct(InvoiceNo)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Mean = round(mean(FrequencyOfSales, na.rm = TRUE), 2),
    SD = round(sd(FrequencyOfSales, na.rm = TRUE), 2),
    Min = round(min(FrequencyOfSales, na.rm = TRUE), 2),
    Max = round(max(FrequencyOfSales, na.rm = TRUE), 2)
  ) %>%
  mutate(Particulars = "Sales frequency to Individual Customers")

# Combine the statistics into a single data frame
combined_stats <- bind_rows(Sales_by_Invoice, TotalSales_customer, sales_frequency_per_customer) %>%
  select(Particulars, everything())  # Ensure Particulars column is at the front

# Create and format the gt table
gt_table <- gt(combined_stats) %>%
  tab_header(
    title = "Descriptive Statistics of Sales") %>%
  cols_label(
    Particulars = "Particulars",
    Count = "Count",
    Mean = "Mean",
    SD = "St.Dev.",
    Min = "Min",
    Max = "Max"
  ) %>%
  cols_align(
    align = "center",
    columns = c("Count", "Mean", "SD", "Min", "Max")
  )

# Display the gt table
print(gt_table)

gt_table %>%
  gtsave(here("results","tables", "gt_table.png"))

```
# Data exploration through figures

Next, density plots were created for Sales across customers and Invoice Values. Both figures are heavily right-skewed.The averages of total sales to a customer and total invoice values appear to be less than 2500 and 1250 sterling pounds respectively supporting the mean values of 2025 and 563 in the descriptive table.

A histogram of purchase frequency of each customer is created by grouping each customer by invoice no.The average frequency appears to be less than 5 times for most of the customers which is also observed in the descriptive statistics table.

```{r}
# Aggregate sales by CustomerID
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  arrange(desc(TotalSales))

#Density Plot of Sales per Customer
# Assuming sales_per_customer is already calculated
p1 <- ggplot(TotalSales_customer, aes(x = TotalSales)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(title = "Density Plot of Sales across Customers", x = "Total Sales", y = "Density")

# Histogram of Sales by Invoice values
## First calculate total invoice value per invoice
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalInvoiceValue = sum(Sales, na.rm = TRUE))

##Create histogram
p2 <- ggplot(Sales_by_Invoice, aes(x = TotalInvoiceValue)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Density Plot of  Invoice Sales",
       x = "Invoice Value", 
       y = "Density") +
  theme_minimal()

# Calculate purchase frequency per customer
purchase_frequency <- mydata %>%
  distinct(CustomerID, InvoiceNo) %>% # Ensure unique InvoiceNo per CustomerID
  count(CustomerID, name = "PurchaseFrequency") # Count the number of invoices per customer
library(ggplot2)

# Histogram of purchase frequencies
p3 <- ggplot(purchase_frequency, aes(x = PurchaseFrequency)) +
  geom_histogram(binwidth = 1, fill = "salmon", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Customer Purchase Frequencies",
       x = "No. of Purchases", 
       y = "Count of Customers") +
  theme_minimal()

# Modify p1 to adjust title font size
p1 <- p1 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p1
# Modify p2 to adjust title font size
p2 <- p2 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p2
# Modify p3 to adjust title font size
p3 <- p3 + theme(plot.title = element_text(size = 11)) # Adjusting font size of the title of p3

plot1<-grid.arrange(p1, p2, p3, ncol = 2, nrow=2)

figure_file = here("results","figures","Sales_distribution.png")
ggsave(filename = figure_file, plot=plot1)

A histogram is also plotted for price range of the products.Most of the products fall below the price of 12.50 sterling pounds.

@@

p4 <- ggplot(mydata, aes(x = UnitPrice)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Unit Price",
       x = "Unit Price", 
       y = "Frequency") +
  theme_minimal()

p4

figure_file = here("results","figures","Price_range.png")
ggsave(filename = figure_file, plot=p4)

```
Next, total purchase by a customer was plotted against purchase frequency.A few outliers were noted.

```{r}
#| message: false
#| warning: false
# Calculate purchase frequency and total invoice value per customer
customer_stats <- mydata %>%
  group_by(CustomerID) %>%
  summarise(PurchaseFrequency = n_distinct(InvoiceNo),
            TotalInvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p5 <- ggplot(customer_stats, aes(x = PurchaseFrequency, y = TotalInvoiceValue)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Purchase Frequency",
       y = "Total Invoice Value") +
  theme_minimal()
plot(p5)

figure_file = here("results","figures","Sales_trend.png")
ggsave(filename = figure_file, plot=p5)

@@

Next, Bar chart is created for Sales Distribution by Country. The top five sales are in Germany, France, Spain, Switzerland and Japan.

### 2024-03-07 products/manuscript/Manuscript.qmd
Project Part 2 - Rev

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

@@

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

Finite mixture models are used to analyze data presumably arising from several separate, yet unobserved, groups. Among these models, Gaussian mixture models and mixture of t-distributions are particularly prevalent. However, such models which are based on symmetric distribution often do not accurately represent data with asymmetric distribution. Skewed data can be more effectively managed using asymmetric distributions, thereby extending their applicability across a broader spectrum of model-based clustering scenarios [1]. If the observed data is positive, distribution is right-skewed and assumed to come from a mixture distribution, employing mixture of gamma density is a rational choice [2].

This study intends to identify distinct customer groups within sales data, with each group comprising customers who exhibit similar characteristics in terms of purchase frequency and amounts. Modeling customer spending as a mixture of gamma distributions helps to identify different spending behaviors. The objective is to develop an Expectation Maximization (EM) algorithm which segregates groups based on their spending patterns, considering the individual purchases transactions made by each customer.

The exploration begins with applying commonly used alternative models to a publicly available data and ultimately applying the gamma mixture model to this data.

@@

@@

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.

The study will also use simulated data to formulate the EM algorithm. Following successful testing, the refined algorithm will be applied to the sales data.

@@

This study aims to answer the question, "How effective is the Expectation Maximization (EM) algorithm in identifying distinct customer groups when accounting for individual spending by each customer, as demonstrated by initial validation using simulated data?"

[1] Lee, S. X., & McLachlan, G. J. (2013). Model-based clustering and classification with non-normal mixture distributions. Statistical Methods & Applications, 22(4), 427-454.

[2] Young, D. S., Chen, X., Hewage, D. C., & Nilo-Poyanco, R. (2019). Finite mixture-of-gamma distributions: estimation, inference, and model-based clustering. Advances in Data Analysis and Classification, 13, 1053-1082.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

@@

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
#knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

@@

@@

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.

The study will also use simulated data to formulate the EM algorithm. Following successful testing, the refined algorithm will be applied to the sales data.

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

@@

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Sales_distribution.png"))
```

@@

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","summarytable.rds"))
knitr::kable(resulttable2)
```

>>>>

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

@@

## General Background Information

This project aims to gain knowledge of customer segmentation from the supervised and un-supervised learning perspectives. In the initial analysis phase, the study applies basic statistical techniques such as logistic regression to analyze a retail store’s data, to grasp general patterns of customer behavior. As the project progresses, it integrates more advanced tool, Expectation Maximization algorithms.

It is crucial that organizations understand their customer’s behaviors and their purchasing patterns. By studying the purchasing patterns of its customers, an organization can design and execute effective customized customer engagement strategies, which can enhance customer service, and consequently increase sales. Thus, in order to understand their customer’s purchasing behavior, organizations can engage in customer segmentation. This analytical approach involves classifying customers into groups exhibiting similar characteristics or purchasing behaviors. When organizations analyze factors, such as purchase frequency and spending habits, they can identify distinct segments within their customer base and subsequently tailor their marketing and sales efforts to meet the unique needs of each group. This tailored personalized approach can enhance customer satisfaction and also improve multiple end results like customer loyalty, operational efficiency, and profitability.

The project will employ a simple model for classification, logistic regression. Since the data doesn't have a labeled categorical variable for classification, a target variable will be created based on the purchase amount and frequency during a fixed period.

The project will also explore the use of mixture models as a potential analytical tool based on the assumption that customer data is inherently complex which can exhibit multiple underlying patterns that are not immediately evident through conventional analysis. Thus, the use of finite mixture models is useful in this project. These models are adept at analyzing data that may contain multiple, distinct groups hidden within the data. Gaussian mixture models and mixtures of t-distributions are examples of the most commonly used finite mixture models in analyzing such data. However, models like these that are primarily grounded in symmetric distributions may not always provide an accurate representation of data characterized by asymmetric distributions. Therefore, in cases where skewed data need to be analyzed, asymmetric models become useful since they offer a broader applicability across various model-based clustering scenarios [@lee2013model]. More specifically, the application of a mixture of gamma densities is a rational choice when dealing with observed data that is positive and exhibits right-skewness, suggesting a mixture distribution origin [@young2019finite]. This project will also explore the application and compare the results of Gaussian mixture model and Gamma mixture model on the retail store’s data.

@@

Initially this data was downloaded from Kaggle.com to assess it's usability for the project. Upon confirming the usability and revisiting the website seeking for further details, it could not be found in Kaggle.com. After few Google searches, the data and details were found on UC Irvine Machine Learning Repository [https://archive.ics.uci.edu/dataset/352/online+retail]. As per the website, this is a transnational data set containing all the transactions occurring between 12/1/2010 and 12/09/2011 for a UK-based and registered non-store online retail. The company primarily sells unique all-occasion gifts and many customers of the company are wholesalers.

@@

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.
(9) Sales: It represents product price multiplied by UnitPrice.

@@

This project aims to gain a comprehensive understanding of customer segmentation from different perspectives; predictive (logistic regression), probabilistic clustering (GMM), and skewness-aware clustering (Gamma Mixture Models).

@@

@@

This project employs sales data , 'superstoredata', of an online retail store based in United Kingdom, to explore purchase patterns of customers outside of United Kingdom.

@@

The data has 541909 observations and nine variables detailed as below:

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.
(9) Sales: It represents product price multiplied by UnitPrice.

One key factor of this data is it accounts sales of a particular item from an Invoice as an observation leading different items sold in the same Invoice to appear as separate observations. As such most of the Invoice numbers appear several times in different observations of the data.

As the purpose of the project is to study purchase pattern of customers residing outside of UK, first all of the UK based observations were dropped from the data which brought down the data size to 46431 observations with 9 variables.

The data inspection pointed out to several issues requiring cleaning such as missing values in CustomerID, negative signs in Quantity and Sales and incorrect format of date. It also pointed out that InvoiceNo, StockCode, Description and Country were coded as character variables in the data which needed to be changed to factors for better data manipulation.

First, there were 1480 missing values in CustomerID. Customer behavior is a crucial part of this analysis which cannot be tracked without the unique customerID. The observations without CustomerID do not help in doing meaningful analysis and are dropped. Further, CustomerID was coded as a numerical variable in the original data and was converted to factor variable during the cleaning process. Country, Description and StockCode were also converted to factors from character variables. Transaction date was also converted to the date format.

There were 1372 observations with negative values in quantity and sales due to order cancellation which were dropped from the data.

A detailed scrutiny of the distribution of UnitPrice revealed that all the high value goods with a Unitprice above 25 were distributed between 75 and 100 percentile. A closer look at this inter-quartile range disclosed that there were 19 categories of product description which had the Unitprice above 25. Some of those categories had descriptions such as CARRIAGE, POSTAGE, Manual, which appeared to be related to logistics. 1216 observations with such descriptions were dropped from the data since those were less likely to be directly related to customer purchase.

The store also sells merchandises in wholesale quantities. A wholesaler potentially purchased in big volumes and in higher frequencies. In absence of clear information on wholesalers, a scatter plot of Total Invoice Value by Purchase Frequency was plotted to identify potential wholesalers. The scatter plot revealed that most of the customers made purchases less than a frequency of 50 during that period and most of the cumulative purchase values were less than 50,000. Assuming that the outliers represented purchases by wholesalers, observations with purchase frequencies above 50 and cumulative Invoice value above 50,000 are dropped from the data. This step excluded 9745 observations from the data. The cleaned data has 32602 observations with 9 variables.

@@

A meaningful analysis of customer spending pattern require the Invoice Sales Amount, total purchase by Customers, frequency of customer purchases during a period or a combination of those. @tbl-resulttable1 is created which provides the descriptive statistics of sales by Invoice value, Customer purchase and frequency of customer purchase. It revealed that there were 1479 transactions with 414 different customers during the period of 2010-12-01 to	2011-12-09. The store recorded a mean sales of GBP 567 per transaction, an average cumulative sales of GBP 2025 to individual customers and sales frequency of above 3 times per customer.

```{r out.width="100%", fig.show='hold', warning=FALSE}
#| label: tbl-resulttable1
#| tbl-cap: "Descriptive Statistics"
#| echo: FALSE
knitr::include_graphics(here("results","tables", "gt_table.png"))

```

@fig-result depicts two density plots for the Cumulative Sales across customers and Invoice Sales and a histogram for Customer purchase frequencies. Both density plots are heavily right-skewed.The averages of total sales to a customer and total invoice values appear to be less than 2500 and 1250 sterling pounds respectively supporting the mean values of 2025 and 563 in the descriptive table. The average frequency appears to be less than 5 times which is also observed in the descriptive statistics table.

```{r}
#| label: fig-result
#| fig-cap: "Distribution of Sales and Purchase frequency"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Sales_distribution.png"))
```

@@

#@fig-result1 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result1
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
#knitr::include_graphics(here("results", "figures", "Sales_distribution.png"))
```

@@

#Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables","summarytable.rds"))
#knitr::kable(resulttable2)
```
